Hydrogen Epoch of Reionization Array (HERA)

### 2016-04-09 manuscript.tex
beginning to reform req and system

\begin{figure}[t]
\centerline{
\includegraphics[height=3in]{plots/wedges.pdf}
}
\caption{The EoR ``window'' and foreground ``wedge'' are shown in this plot, which splits the wavenumber into $\kvpr$ (horizontal) and $\kpar$ (vertical) components.
Smooth-spectrum foregrounds with an intrinsic spectrum indicated by the gray bar along the bottom create the ``wedge'' when they interact with the interferometer's chromatic response, shown with a foreground image from the MWA.  Systematics can extend the wedge to the upper left.  The EoR window (shown with a simulation
from \cite{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
\label{fig:wedge}
\end{figure}

@@

Since the thermal noise per visibility baseline may be expressed as
\begin{equation}
V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
\label{eq:sensitivity_per_baseline}
\end{equation}
where $\tau$ is the integration time (see {\em e.g.} TMS),
to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq. \ref{eq:Pk_baseline}.  Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
\begin{equation}
\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
\label{eq:sensitivity}
\end{equation}
where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
(see \cite{parsons_et_al2012b}).

Fig. \ref{fig:boost} (left) plots this ``redundancy boost'' factor for various configuration types relative to an array in a 19-element hexagonal configuration.  They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).  Note that the (a) hex is based on the configuration used for HERA (14.6m center-to-center spacing); (b) grid for that of PAPER (4m N-S, 16m E-W spacing); (c) Imaging-128 for MWA \citep{tingay_et_al2013}; and (d) Imaging-48 for LOFAR \citep{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).  The plot shows that for a fixed number using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.  Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).  This advantage does not extend to the other approaches discussed below and imaging configurations would be preferred.  Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are includes.

Another approach to improving sensitivity is to make the collecting area per element larger.  In the limit, this yields a fully filled aperture as mentioned.  A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' pass passes within its field.  One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.  This adds additional complexity and potential systematics, however it is the method employed by both MWA and LOFAR.  Fig. \ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19.  The circles near the array name indicate the relative sizes of the element.  Note that MWA and LOFAR are cross-hatched to indicated phased array tiles.

@@

In Fig. \ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $|\kvpr|$, and green lines denote total magnitude ($|{\kvec}|$).  The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).  Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer really appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
Also in Fig. \ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first instance.

@@

\subsection{Antenna Design and Configuration}
The primary new feature of HERA over previous generation experiments (in this context, primarily PAPER) is the use of large elements to increase the sensitivity.  The cost-performance optimization will 
be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge, to open the widest possible window for EoR measurements, 
HERA must use close-packed antennas that
minimize signal reflections over long delays and deliver significant forward gain relative to their horizon response.
Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
indicate that a moderately large parabolic dish with a short focal height can meet these requirements
\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.  Note that to pursue map-making and hybrid techniques, additional antenna outriggers are valuable additions, so the design
should be able to accommodate ~1km baselines.  In referring to the EOR sensitivity of the array however, the number of antennas just in the core is sometimes used.

@@

From these measurements and models, HERA is required to cover a redshift range of $z=$6 to 13, corresponding to a frequency range of 101 - 203 MHz.  We adopt 100-200 MHz for full performance as the requirement.  Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.  The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.  The 100 MHz bandwidth is also the minimum fully digitized and correlated bandwidth.

@@

\subsection{Sensitivity Optimization}
\label{sec:cost}
As an instrument to characterize the power spectrum over its evolution, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-13, along with a very robust detection at the peak.  As seen in Fig. \ref{fig:Sensitivities} and Table \ref{tab:signif}, an array comprising 320 14-m core antennas (note, labelled HERA-350, which includes the outriggers) has sensitivity to a
fiducial reionization model across that redshift range.  We therefore adopt the delay-spectrum power spectrum sensitivity of 320 14-m close-packed antennas as the minimum sensitivity needed.  
For this sensitivity, the number of elements needed for a given diameter is
\begin{equation}
N_{\text core} = (320*14)/D
\end{equation}
where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.  Note that the number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration.  This dependency differs from Mellema and analytical epressions, mostly likely due to the assumed configuration.
Optimization is 
therefore done for this sensitivity against cost for various diameters and total element numbers, while trying to minimize multiple reflections on long delays.  Figure \ref{fig:costfig} summarizes this optimization, which
is discussed below.

To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity at the minimum discussed in the previous paragraph.  This included the full system, except the post-processing (that is, everything after the archived correlator output).  
Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
\begin{figure}[t]
\centerline{
\includegraphics[width=7.5cm]{plots/costfigNew.png} 
}
\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).  The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.  SEE TEXT...
\label{fig:costfig}}
\end{figure}

@@

\noindent As described in the previous section, we have applied 
critical insights from first generation 21\,cm EoR experiments to define
the requirements for HERA---an instrument that ensures foregrounds remain bounded
within the wedge while \emph{delivering the sensitivity for
high-significance detections of the 21\,cm reionization power spectrum with
established foreground filtering techniques} 
\citep{pober_et_al2014,greig_and_mesinger2015}.
In this section, we summarize key features of the HERA design (see Table~\ref{tab:BasicParameters}) 
and system architecture (see Fig.~\ref{fig:overallBlockDiagram}).
This architecture
directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
analog, digital, and real-time processing systems deployed for PAPER-128.  As HERA develops, this
architecture is incrementally upgraded to improve performance and add features while simultaneously
addressing issues of modularity and scalability.  As with PAPER, HERA proceeds in stages of development,
with annual observing campaigns driving a cycle of development, testing, system integration, calibration, and analysis.
This cycle ensures that HERA's instrument is always growing, that systematics are being found and eliminated at the
earliest build-out stages, that data analysis pipelines are tested and debugged while data volumes are smaller,
and that HERA is always producing high quality science.

@@

\subsection{Antenna Element}
\label{sec:antenna}
\vspace{-5pt}

@@

The construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.  A close-up of one antenna is shown in Fig \ref{fig:elementpic}.  The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.

The center of the antenna is a cast concrete donut-shaped hub with 12 angled sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.  The twelve spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.  Note that an ideal moment-loaded beam actually attains a true parabola.  Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.

The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.  If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).  Therefore, at about 2.1 meters another horizontal member is placed, which launches another parabolic spar.  The simple Ruze loss is then less than 1\%, which is much more acceptable.  Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.  The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable bridge).  The additional intermediate spars also make the panel sizes much more manageable and better secured.  The panel arrangement is also set by a standard width of mesh roll of 1.22m.

@@

The feed is to be deployed along the two separate phases.  Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.  The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on dish), standing waves (blockage, {\em e.g.} Baars), and frequency response.  The design parameters were size of backplane, height of mast, height of skirt, and cone versus cylinder.  Cones were ruled out since it was found they introduce additional frequency structure with no real improvement in performance and a range of cylinders were evaluated.  The phase 1 design is documented in a memo \cite{feedmemo}, while additional studies are on-going at NRAO, University of Cambridge and Stellenbosch University.  The chosen feed design is shown in Figure X.

Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.

As mentioned, the feed is supported by lines from the three poles via a rigid support to which the feed cage and dipole attach.  Another line goes down centrally to the hub to hold the feed rigidly in place at the correct height.

The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).  The 150m length of 75 $\Omega$ cable also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.

\noindent The novel design of HERA's antenna element (Figs. \ref{fig:HERApictures} and \ref{fig:orbcommexptandbeammap}), now extensively field-tested with MSIP seed funding, 
is one of the critical advances that enables HERA to achieve its science
goals cost-effectively.  The 14-m fixed zenith-pointing parabolic dish strikes an optimal
balance between sensitivity and systematics.  The large collecting area of a
HERA element yields nearly 5 times the sensitivity of an MWA tile and more
than 20 times that of a PAPER element, but it does so without substantially
degrading our ability to isolate and remove foreground emission on the basis of
spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
\Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
used to measure reionization.
To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.

@@

With observing wavelengths of 1.5 to 3 m, meeting the required accuracy
in placing the components of HERA's dish is straightforward.
Dish centers are fixed
by a concrete hub placed with final surveyed accuracy of $\sim$10 cm.  These
hubs constrain radial PVC spars, tensioned into approximate parabolas against a
rim supported by utility poles.  The resulting faceted paraboloid has
a measured RMS surface accuracy of 3 cm, which is well within the 0.1$\lambda$ tolerance for
coherent phasing.  Feed placement relative to the dish is the most stringent requirement, with
a tolerance of 5 cm.  This placement is ensured by spring-tensioned Kevlar lines attached to
three utility poles at the dish rim and a 165-lb tensioning down to the concrete hub with
a fixed length line that ensures a correct focal height.  With 19 dishes constructed in South Africa
and 2 in Green Bank, we are confident in the ability of our trained field teams to construct HERA
elements to specification.

@@

\noindent HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
The core is supplemented by 30 additional outrigger elements 
to tile the $uv$-plane with instantaneously complete sub-aperture 
sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
direction-dependent antenna responses \citep{dillon_parsons2016}.
Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
of the residual fractional noise per antenna after averaging.

@@

{\bf Karoo Array Processing Building (KAPB)}.
The KAPB houses the switch and CPUs
that 
complete the HERA correlator system.  A fiber optic bundle enters the KAPB and patches
into local fiber optic cables. These cables terminate in optical transceivers which plug into a 240-port 10 GbE switch.
Such switches, while large, are readily available commercially today.  Also connected to
this switch are 30 servers, each hosting two dual-GPU graphics cards and two dual
10 GbE network interface cards, which implement the cross-multiplication (X-Engine) component
of the correlator during observations.  This estimate for the number of X-Engine servers
is extrapolated from current GPU servers deployed on PAPER, assuming no improvement in bus
speeds for transferring data into the GPU cores, but assuming that the computational
capacity of GPU cores doubles once according to Moore's Law prior to purchasing
these servers in year three of the project.
Output data from the correlator are written to the data storage system described
in the following section.

>>>>

\begin{figure}[t]
\centerline{
\includegraphics[height=3in]{plots/wedges.pdf}
}
\caption{The EoR ``window'' and foreground ``wedge'' are shown in this plot, which splits the wavenumber into $\kvpr$ (horizontal) and $\kpar$ (vertical) components.
Smooth-spectrum foregrounds with an intrinsic spectrum indicated by the gray bar along the bottom create the ``wedge'' when they interact with the interferometer's chromatic response, shown with a foreground image from the MWA.  Systematics can extend the wedge to the upper left.  The EoR window (shown with a simulation
from \citealt{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
\label{fig:wedge}
\end{figure}

@@

Since the thermal noise per visibility baseline may be expressed as
\begin{equation}
V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
\label{eq:sensitivity_per_baseline}
\end{equation}
where $\tau$ is the integration time (see {\em e.g.} TMS),
to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq. \ref{eq:Pk_baseline}.  Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
\begin{equation}
\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
\label{eq:sensitivity}
\end{equation}
where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement (see \citealt{parsons_et_al2012b}).  Dependencies on redshift and wavenumber have been included.

Fig. \ref{fig:boost} (left) plots this ``redundancy boost'' factor for various configuration types relative to an array in a 19-element hexagonal configuration.  They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).  Note that (a) {\em hex} is based on the configuration used for HERA (14.6m center-to-center spacing); (b) {\em grid} for that of PAPER (4m N-S, 16m E-W spacing); (c) {\em imaging-128} for MWA \citep{tingay_et_al2013}; and (d) {\em imaging-48} for LOFAR (\citealt{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).  The plot shows that for a fixed number of elements using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.  Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).  This advantage does not extend to the other approaches discussed below and imaging configurations would be preferred.  Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are included.

Another approach to improving sensitivity is to make the collecting area per element larger.  In the limit, this yields a fully filled aperture as mentioned.  A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' sky passes within its field-of-view.  One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.  This adds additional complexity and potential systematics, however it is the method employed by both MWA and LOFAR.  Fig. \ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 1/2.  The circles near the array name indicate the relative sizes of the element.  Note that MWA and LOFAR are cross-hatched to indicated phased array tiles.

@@

In Fig. \ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $\kpr$, and green lines denote total magnitude ($|{\kvec}|$).  The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).  Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
Also in Fig. \ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first instance.

@@

\subsection{General Antenna Design and Configuration}
The primary new feature of HERA over previous generation experiments (in this context, primarily PAPER) is the use of large elements to increase the sensitivity, as indicated in Fig. \ref{fig:boost}.  The cost-performance optimization will 
be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge, to open the widest possible window for EoR measurements, 
HERA must use close-packed antennas that
minimize signal reflections over long delays and deliver significant forward gain relative to their horizon response.
Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
indicate that a moderately large parabolic dish with a short focal height can meet these requirements
\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.  Note that to pursue map-making and hybrid techniques, additional antenna outriggers are valuable additions, so the design
should be able to accommodate $\approx$1km baselines.  Note that in referring to the EOR sensitivity of the array, often the number of antennas just in the core are
indicated -- that is for EOR HERA-320 and HERA-350 are often used interchangeably.

@@

From these measurements and models, HERA is required to cover a redshift range of $z=$6 to 13, corresponding to a frequency range of 101 - 203 MHz.  We adopt 100-200 MHz for full performance as the requirement.  Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.  The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.

@@

\subsection{Delay Response}
\label{sec:delayspec}
Need to include initial 60 at 60 and Nithya etc etc etc

\subsection{Sensitivity Optimization}
\label{sec:cost}
As an instrument to characterize the power spectrum over the evolution of the EoR, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-13, along with a very robust detection at the peak.  As seen in Fig. \ref{fig:Sensitivities} and Table \ref{tab:signif}, an array comprising 320 14-m core antennas (note, labelled HERA-350, which includes the outriggers) has sensitivity to a
fiducial reionization model across that redshift range.  We therefore adopt the delay-spectrum power spectrum sensitivity of 320 14-m close-packed antennas as the minimum sensitivity needed.  
For this sensitivity, the number of elements needed for a given diameter is
\begin{equation}
\label{eq:ND}
N_{\text core} = (320*14)/D
\end{equation}
where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.  Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration for wavenumbers greater than 0.15.  This dependency differs from \cite{2013ExA....36..235M} and analytical expressions, mostly likely due to the assumed configuration.
Optimization is 
therefore done for this sensitivity against cost for various diameters and total element numbers, while trying to minimize multiple reflections on long delays.  Figure \ref{fig:costfig} summarizes this optimization, which
is discussed below.

To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, with the total number set by Eq. \ref{eq:ND}.  This included the full system, except the post-processing (that is, everything after the archived correlator output).  
Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.

OTHER LINES IN FIG 12.

\begin{figure}[t]
\centerline{
\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
}
\caption{\small The background coloring with colorbar shows the relative cost/performance at a fixed performance as a function of element diameter.  Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).  The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.  SEE TEXT...
\label{fig:costfig}}
\end{figure}

@@

As described in the previous section, we have applied 
critical insights from first generation 21\,cm EoR experiments to define
the requirements for HERA---an instrument that ensures foregrounds remain bounded
within the wedge while delivering the sensitivity for
high-significance detections of the 21\,cm reionization power spectrum with
established foreground filtering techniques
\citep{pober_et_al2014,greig_and_mesinger2015}.
In this section, we summarize key features of the HERA design (see Table~\ref{tab:BasicParameters}) 
and system architecture (see Fig.~\ref{fig:overallBlockDiagram}).
This architecture
directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
analog, digital, and real-time processing systems deployed for PAPER-128.  As HERA develops, this
architecture is incrementally upgraded to improve performance and add features while simultaneously
addressing issues of modularity and scalability.  As with PAPER, HERA proceeds in stages of development,
with annual observing campaigns driving a cycle of development, testing, system integration, calibration, and analysis.
This cycle ensures that HERA's instrument is always growing, that systematics are being found and eliminated at the
earliest build-out stages, that data analysis pipelines are tested and debugged while data volumes are smaller,
and that HERA is always producing high quality science.

For details of the first generation HERA system (the existing and commissioned PAPER signal path), see \cite{PAPERSYS}.  This paper will discuss the new system to be deployed beginning and 2016 and tested in conjunction with the previous system.  Fig \ref{fig:overallBlockDiagram} provides a system block diagram of the new version.

@@

\subsection{Antenna}
\label{sec:antenna}
\vspace{-5pt}

@@

\subsubsection{Element Construction}
To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.   The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.  Two of the three posts are shared with adjacent antennas.

The center of the antenna is a cast concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.  The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.  Note that an ideal moment-loaded beam actually attains a true parabola.  Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.  The PVC essentially acts as a spatial filter.

The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.  If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).  Therefore, at about 2.1 meters another horizontal member is placed, which launches another parabolic spar.  The simple Ruze loss is then less than 1\%, which is much more acceptable.  Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.  The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).  The additional intermediate spars also make the panel sizes much more manageable and better secured.  The panel arrangement is also set by a standard width of mesh roll of 1.22m.

@@

\subsubsection{Feed}
The feed is to be deployed along the two separate phases.  Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.  The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on dish), standing waves (blockage, {\em e.g.} Baars), and frequency response.  The design parameters were size of backplane, height of mast, height of skirt, and cone versus cylinder.  Cones were ruled out since it was found they introduce additional frequency structure with no real improvement in performance and a range of cylinders were evaluated.  The phase 1 design is documented in a memo \cite{feedmemo}, while additional studies are on-going at NRAO and Cambridge University.  The chosen feed design is shown in Figure X.

%Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  
%Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  
%Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.

The feed is supported by lines from the three poles via a rigid support to which the feed cage and dipole attach.  Another line goes down centrally to the hub to hold the feed rigidly in place at the correct height.

The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).  The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.  This will be replaced by the 35m cable/node design.

The novel design of HERA's antenna element (Figs. \ref{fig:HERApictures} and \ref{fig:orbcommexptandbeammap}), now extensively field-tested with MSIP seed funding, 
is one of the critical advances that enables HERA to achieve its science
goals cost-effectively.  The 14-m fixed zenith-pointing parabolic dish strikes an optimal
balance between sensitivity and systematics.  The large collecting area of a
HERA element yields nearly 5 times the sensitivity of an MWA tile and more
than 20 times that of a PAPER element, but it does so without substantially
degrading our ability to isolate and remove foreground emission on the basis of
spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
\Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
used to measure reionization.
To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.

@@

With observing wavelengths of 1.5 to 3 m, meeting the required accuracy
in placing the components of HERA's dish is straightforward.
Dish centers are fixed
by a concrete hub placed with final surveyed accuracy of $\sim$10 cm.  These
hubs constrain radial PVC spars, tensioned into approximate parabolas against a
rim supported by utility poles.  The resulting faceted paraboloid has
a measured RMS surface accuracy of 3 cm, which is well within the 0.1$\lambda$ tolerance for
coherent phasing.  Feed placement relative to the dish is the most stringent requirement, with
a tolerance of 5 cm.  This placement is ensured by spring-tensioned Kevlar lines attached to
three utility poles at the dish rim and a 165-lb tensioning down to the concrete hub with
a fixed length line that ensures a correct focal height.  With 19 dishes constructed in South Africa
and 2 in Green Bank, we are confident in the ability of our trained field teams to construct HERA
elements to specification.  CAMBRIDGE

@@

\noindent HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
The core is supplemented by 30 additional outrigger elements 
to tile the $uv$-plane with instantaneously complete sub-aperture 
sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
direction-dependent antenna responses \citep{dillon_parsons2016}.
Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
of the residual fractional noise per antenna after averaging.  DILLON REF

@@

\subsection{Processing}
\label{sec:processing}
{\bf Karoo Array Processing Building (KAPB)}.
The KAPB houses the switch and CPUs
that 
complete the HERA correlator system.  A fiber optic bundle enters the KAPB and patches
into local fiber optic cables. These cables terminate in optical transceivers which plug into a 240-port 10 GbE switch.
Such switches, while large, are readily available commercially today.  Also connected to
this switch are 30 servers, each hosting two dual-GPU graphics cards and two dual
10 GbE network interface cards, which implement the cross-multiplication (X-Engine) component
of the correlator during observations.  This estimate for the number of X-Engine servers
is extrapolated from current GPU servers deployed on PAPER, assuming no improvement in bus
speeds for transferring data into the GPU cores, but assuming that the computational
capacity of GPU cores doubles once according to Moore's Law prior to purchasing
these servers in year three of the project.
Output data from the correlator are written to the data storage system described
in the following section.
